BoostEM

### 2014-05-09 paper/paper.tex
some more bs

It is well established in Machine Learning that labels are difficult to obtain while unlabeled data are abundant in many situations where we would like to classify new instances. Examples of such situations are labeling x-rays as benign or malignant, tweets as instances of bullying, news articles as interesting to a given reader, or whether a photograph contains a given object. Because of this scarcity of labeled instances, using semi-supervised learning techniques, such as Expectation Maximization, to combine unlabeled and labeled data into an integrated model is a good idea to potentially improve learner accuracy and reduce human labour.

@@

\section{Background}
\subsection{Boost}

~\cite{Freund:1997:DGO:261540.261549}

@@

\EndFor
\State \textbf{return} $\arg \max_{y} \sum_{i=1}^{T}\bigg(\log(\frac{1}{\beta_{t}}\bigg) \delta(\bf{x},y)$
\EndProcedure
\end{algorithmic}
\end{algorithm}

By using a weighted EM, we can place more emphasis on the instances our learner incorrectly classifies in the Gaussian distribution of instances we assume. Or more correctly, since there is no ground truth to the unlabeled instances, we can penalize every prediction from learner $H$ on unlabeled instance $x$ by $1-p(H(x))$, where $p(\dot)$ is the distribution of labels from EM.  In this way, the learner penalizes unexpected labels (ones with small probability) more than expected ones.

@@

@@

\section{Conclusions}
%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

>>>>

In many machine learning classification problems, labels are difficult to obtain while unlabeled data are abundant. Examples of such situations are labeling x-rays as benign or malignant, tweets as instances of bullying, news articles as interesting to a given reader, or whether a photograph contains a given object. Because of this scarcity of labeled instances, using semi-supervised learning techniques, such as Expectation Maximization, to combine unlabeled and labeled data into an integrated model is a good idea to potentially improve learner accuracy and reduce human labour.

@@

\section{Background}
\subsection{Boost}
Boosting is an ensemble method which uses a set of learners trained on weighted versions of a training set. A popular boosting algorithm is known as AdaBoost (short for Adaptive Boost) originally detailed in ~\cite{Freund:1997:DGO:261540.261549}.  AdaBoost works by giving each training set instance a weight $W_{i}$; initially all weights are equal. Over some number of iterations, AdaBoost trains a base learner (ID3 decision trees, Naive Bayes, etc) proportionally to the weights of each instance. The weighted training set error $\epsilon$ of this base learner is calculated and if more than half of the instances by weight were incorrect AdaBoost quits.  Otherwise, $\beta=\frac{\epsilon}{1-\epsilon}$ is applied to all of the correctly guessed instances, decreasing their weights.  This means that in the next iteration, incorrectly classified instances will be given more importance.  AdaBoost then essentially drives up the importance of difficult to classify training set instances in an effort to correctly learn them.  To classify a new instance $\bf{x}$, with base learners $H_{1} \dots H_{T}$  $\arg \max_{y} \sum_{i=1}^{T}\bigg(\log(\frac{1}{\beta_{t}})\bigg) \delta(H_{t}(\bf{x}),y)$ where $\delta(a,b)$ is the indicator function.

Empirical studies ~\cite{dietterich1997machine} have shown that in a large majority of instances AdaBoost improves accuracy over the base learner.  There are a few occasions however where using Boosting can hurt performance.  This implies that when using AdaBoost, one should take care to compare results from the combined learner to those from a base learner.

@@

\EndFor
\State \textbf{return} $\arg \max_{y} \sum_{i=1}^{T}\bigg(\log(\frac{1}{\beta_{t}})\bigg) \delta(H_{t}(\bf{x}),y)$
\EndProcedure
\end{algorithmic}
\end{algorithm}

 Using an unweighted version of Expectation Maximization is possible, but since the training set data itself is not changing, the fractional counts will not change from iteration to iteration. By combining the AdaBoost weights with the EM algorithm, BoostEM is able to alter the distribution and get an updated set of fractional counts in each iteration of BoostEM.   By using a weighted EM, we can place more emphasis on the instances our learner incorrectly classifies in the Gaussian distribution of instances we assume. Or more correctly, since there is no ground truth to the unlabeled instances, we can penalize every prediction from learner $H$ on unlabeled instance $x$ by $1-p(H(x))$, where $p$ is the distribution of labels from EM.  In this way, the learner penalizes unexpected labels (ones with small probability) more than expected ones.

@@

\subsection{Implementation Details}
In our testing, we only attempt to test BoostEM in classification.  Because of this, we choose to use a Gaussian mixture model as the underlying model for data. Given a classification label, we assume the conditional distribution of the data to be gaussian distributed. One of the benefits of assuming an underlying distribution in Expectation Maximization is that it allows us to use both labeled data and a potentially much larger amount of unlabeled data in the maximization step. This choice may not be appropriate for all data sets; we choose Gaussian mixture models because many data sets can be roughly approximated by such a model.  A different underlying models which provides a better fit to the data will produce likely produce better results.  An overview of EM and Gaussian Mixture Models can be found in ~\cite{zhu2009introduction}.

@@

\section{Conclusions}
We present BoostEM, a semi-supervised ensemble method.  BoostEM works by combining fractional counts from using the Expectation Maximization algorithm with the AdaBoost ensemble method.  We tested BoostEM on four large datasets from the UCI repository, performing 10-fold cross validation and five random restarts for EM.  We compared using 20\% labeled data to 5\% labeled data, the rest being unalbeled and used ID3 decision trees and Naive Bayes as base learners.  BoostEM was compared to Boosting with 100\% of the data, Boosting with only the percentage of labeled data, performing only Expectation Maximization on the unlabeled data and using the base learner, using the base learner on all the data, and finally using only the subset of unlabeled data with a base learner. We found that in the majority of settings, BoostEM outperforms using only Boosting or only Expectation Maximization.  Additionally, we found that across data sets and base learners that BoostEM performs comparably to only using a base learner, but with only one fifth of the labeled data. Since with specific parameters, BoostEM can replicate both AdaBoost and EM, BoostEM can always achieve performance on a data set at least as well as them, and we have shown that in many cases BoostEM can also improve performance.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
Zach would like to thank his cats, who only broke one of his coffee mugs in the course of this project.
